---
title: "Assignment 4"
author: "Caroline He"
date: "11/16/2021"
output: 
  github_document:
    html_preview: false
  html_document: default
always_allow_html: true
---

## R packages
```{r}
library(dplyr)
library(parallel)
```


# HPC

## Problem 1: Make sure your code is nice

Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}
# rewrite 
fun1alt <- function(mat) {
  ans <- rowSums(mat) 
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}
# rewrite
fun2alt <- function(mat) {
  ans <- t(apply(mat,1,cumsum))
  ans
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)

```

## Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():

```{r}
system.time({
  cl <- makePSOCKcluster(4L)
  clusterSetRNGStream(cl, 1231) 
  clusterExport(cl, "sim_pi")
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n = 10000))
  print(mean(ans))
  stopCluster(cl)
  ans
})

```

# SQL

Setup a temporary database by running the following chunk

```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

When you write a new chunk, remember to replace the r with sql, connection=con. Some of these questions will reqruire you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp

## Question 1

How many many movies is there avaliable in each rating catagory.

```{sql, connection = con}
SELECT rating,
  COUNT(*) AS count
FROM film
GROUP BY rating
```

# Question 2

What is the average replacement cost and rental rate for each rating category.

```{sql, connection = con}
SELECT
  rating,
  AVG(replacement_cost) AS avg_replacement_cost, 
  AVG(rental_rate) AS avg_rental_rate
FROM film
GROUP BY rating
```

# Question 3

Use table film_category together with film to find the how many films there are with each category ID

```{sql, connection = con}
SELECT 
COUNT(*) AS "number of films", 
category_id
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
GROUP BY category_id
```


# Question 4

Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{sql, connection = con}
SELECT 
COUNT(*) as "number_of_films", 
name
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
INNER JOIN category
ON category.category_id = film_category.category_id 
GROUP BY name
ORDER BY number_of_films DESC
```





